title

### 2025-02-28 docs/index.qmd
scripts

---
title: title
subtitle: subtitle
shorttitle: A Short title # aog-article-pdf
shortauthors: Isaiah      # aog-article-pdf
authors:
  - name: Isaiah Espinoza
    email: gespinoz@umd.edu
    roles: author
    affiliation:
      - name: University of Maryland
        department: Department of Government and Politics
        city: College Park
        state: MD
        country: US
reference-section-title: References
date: "2024-10-16"
date-modified: today
crossref:
  custom: 
    - key: hyp
      kind: float
      reference-prefix: H
      latex-env: hypothesis # required for PDF format
---

@@

# load packages
library(tidyverse)

# load data
load("data/research-paper-data-20241029.Rdata")

# load data dictionary/codebook
codebook <- read.csv("data/survey_codebook.csv")
data_dict <- read.csv("data/data_dictionary.csv")

# also load the raw data dict for comparison
raw_spss_dict <- read.csv("data/raw_spss_dict.csv")
```

```{r}
#| label: post_processing
#| include: false

# Some post-processing of the dataframe :::::::::::::::::::::::::::::::::::::###

# recode factor levels 
data <- data |> 
  # use `pewmethods` to preserve order of factor levels
  mutate(across(c(q19, q20, q21, q22, q23,q24, q25, q26,
                  q30, q31, q32, q33, q34, q35, q36, q37), 
                ~pewmethods::fct_case_when(
    .x == "Not at all confident" |
      .x == "Not at all committed" ~ "Not at all",
    .x == "Not too confident" |
      .x == "Not too committed"|
      .x == "Not too concerned"|
      .x == "Somewhat unconcerned" ~ "Not that much",
    .x == "Somewhat confident" |
      .x == "Somewhat committed" ~ "Somewhat",
    .x == "Very confident" |
      .x == "Very committed" ~ "Very",
    TRUE ~ NA
  )))

# re-assign var labels to variables
# To quickly assign the variable labels, first create a named vector via
# deframe() with values as the variable labels and names as the variable names.
data_labels <- data_dict |>
  select(var, label_var) |> 
  deframe()

# Now assign the labels using the splice operator. Using the splice operator,
# labels are assigned via matching against the variable name, which means that
# variable order does not matter.
data <- data |> 
  labelled::set_variable_labels(!!!data_labels)

```{r}
#| label: fns
#| include: false

source(here::here('src', 'funs.R'))

@@

That being said, the general public perception, attitudes, or even stereotypes about military veterans are significant to consider. Although military recruitment shows a downward trend as of late, public perception of veterans are overwhelmingly positive [@kleykamp2023]. Veterans are a particularly potent group where mention of veteran status seems to have a positive, calming, or nullifying effect on attitudes. For instance, recent research shows that, during his campaign in the 2020 Democratic Primaries, Pete Buttigieg's military background mitigated discrimination against him when he was presented as a veteran married to man [@magni2024]. Similar research has also found that a candidate's veteran status affords them better evaluations regarding competency in particular issue areas (e.g., war competence) [@hardy2019]. Moreover, veteran status seems to mitigate or nullify usual stigmas associated with mental illness. That is to say, there is negative stigma associated with mental illness [@corrigan2002] and such stigma incurs labor market discrimination [@hipes2016], but evidence suggests that veteran status overrides such stigma and discrimination [@maclean2014]. Or, in another light, mental illness is seemingly more *understandable* (i.e., permissible) for veterans given the presumptive reasons for their mental strife. And media framing as such plays an important role on public perception [@kleykamp2015].

## Design and Survey Measures

To test the theory that publicized efforts to recruit veterans to work as election staff and volunteers would improve trust in elections and ease election anxieties, a recent survey experiment was embedded in a survey developed and conducted by the Center for Democracy and Civic Engagement (CDCE) at the University of Maryland. The survey was fielded from August 29th, 2024 to September 18th, 2024 on a non-probability sample of 1,287 U.S. citizens 18 years of age or older. Respondents were randomly split into either the treatment (n = 650) control (n = 637) conditions[^4].

Survey participants read either a treatment or control vignette, which was a fabricated news article about efforts in Maricopa County, AZ to recruit election staff and volunteers for the 2024 general election. The treatment vignette referred to a program designed to recruit veterans and their family members and describes an interviewee "Jordan Braxton" as an Army veteran. The control vignette simply omitted any mention of veterans and their family members, and didn't describe "Jordan Braxton" as an Army veteran. Beyond those small differences and the headlines, the article vignettes are identical. Any effects can be attributed to the information about veterans in the treatment vignette[^5].

It should be noted that Maricopa County, AZ was chosen as the setting of the story in the vignettes due to the increased scrutiny levied toward election administration there after the 2020 election [@giles2021; @maricopacountyelectionsdepartment2022]. A block of survey items asked specifically about Maricopa County, AZ, followed by an identical block of items that asked the same questions about one's local area. Results between the items where the setting differs are compared and discussed.

### Trust in Elections

@@

Note that, except for the item assessing voter confidence (i.e., votes counted as voters intend), most other question items from the 2022 SPAE inquire into the voter experience after the election has occurred. The items in this study, however, all inquire into confidence over one's expectations in anticipation of upcoming elections. In addition, other survey questionnaires ask similar questions that specifically contrast between a respondent's own local community, their state, and nationwide. This is in addition, or complementary to, items that distinguish between local area officials, state election officials, state government, and the U.S. federal government. The survey of this study, instead, focused on attitudes regarding election workers such as election officials, staff, and volunteers. As such, the items are utilized to comprise a component[^6] I broadly refer to as *trust in elections*.

A set of five survey items measured public trust in elections by inquiring into the degree of confidence that votes will be counted as voters intend, confidence that electoral systems are secure from technological threats, perceived commitment of election staff, confidence that outcomes will be fair, as well as confidence that the voting process will be fair. The items measuring trust and confidence in elections include:

::: {#tbl-1}
| **Trust in Elections** |
|---|
| How confident are you that votes in Maricopa County, AZ will be counted as voters intend in the elections this November?[^7] |
| How confident are you that election officials, their staff, and volunteers in Maricopa County, AZ will do a good job conducting the elections this November? |
| Think about the election staff and volunteers who handle the administration and conduct of elections in Maricopa County, AZ. How committed do you think they will be to making sure the elections held this November are fair and accurate? |
| How confident are you that the voting process will be fair in Maricopa County, AZ? |
| How confident are you that election systems in Maricopa County, AZ will be secure from hacking and other technological threats? |
:::

Each item in this series presented four response options from "Not at all [confident/committed]", "Not too [confident/committed]", "Somewhat [confident/committed]", and "Very [confident/committed]". In total, six survey items measure different aspects of trust and confidence in election administration. From these items, I generated a composite score for each respondent that represents their variable level of trust in elections.

### Expecatation of Electoral Fraud

A series of five survey items captured *expectation of electoral fraud*. These items were prefaced with the question, "How likely do you think any or all of the following will happen during this year´s elections in Maricopa County, AZ?" Each item presented four response options: "Not likely at all", "Not too likely", "Somewhat likely", and "Very likely".

::: {#tbl-2}
| **Expectation of Electoral Fraud** |
|---|
| There will be voter fraud, that is, people who are not eligible to vote will vote, or vote more than once |
| Many votes will not actually be counted |
| Many people will show up to vote and be told they are not eligible |
| A foreign country will tamper with the votes cast in this area to change the results |
| Election officials in Maricopa County, Arizona will try to discourage some people from voting |
:::

Similar to the items capturing trust in elections (@tbl-1), I computed a composite score for each respondent reflecting their expectations of electoral fraud.

@@

The outcome of interest concerns both public trust in election administration and the expectation of electoral fraud. Because these variables are assessed by multiple Likert items, conventional methods call for combining the items of each set into separate Likert scales, also known as a summated rating scales.
However, many of the items comprising the trust scale directly oppose items capturing expectations of electoral fraud. For instance, one item has the respondent report their confidence that votes will be counted as voter's intend; whereas a contrasting item has the the respondent report their expectation that many votes will not be counted. These two items directly oppose each other, but there's no guarantee that individual responded as such.

Since many of the items are not symmetric---i.e., response options are not bipolar with contrasting positive and negative response options---then a "trust in elections" scale could only be understood as a measure of trust in the positive use of the term (i.e., unipolar). Likewise, the expectation of fraud items fit to a scale that corresponds to a unipolar measure of *confident distrust* so to speak.

I would expect composite sum or mean scores, and item responses, on these two measures to be inversely correlated among the sample, but not mutually exclusive. Indeed, this is what I find. Upon construction of the two multi-item composite scales of *trust in elections* and *expectation of fraud*, internal consistency reliability coefficient $\alpha$ were estimated for both and found to be adequate ($\alpha$ > 0.8); the trust in elections scale $\alpha$ = 0.93 and the expectation of electoral fraud scale $\alpha$ = 0.86. In other words, the items in each of the scales were intercorrelated well enough to compose a single scale of a (presumed) unidimensional construct. I then plotted separate histograms of scores across the sample to visualize the distribution for each seen in @fig-dist. Note that for both scales, higher values indicate less expectation of electoral fraud and greater trust in elections, respectively.

```{r}
#| label: fig-dist
#| fig-cap: "Histograms of Composite (mean) Scores for Trust in Elections And Expectation of Electoral Fraud"

# trust scores are not normally distributed, bi-modal, neg skew
trst.scores.hist <- ggplot(data, aes(trst.mean.scores))+
  geom_histogram(aes(y = after_stat(density)), bins = 15,
                 color = 1, fill = "grey")+
  geom_density(lwd = 1.5, color = 1, fill = 1, alpha = 0.35)+
  geom_segment(x = round(mean(data$trst.mean.scores), 2), y = 0, yend = 0.80, color = "red")+
  labs(
    x = "Trust in Elections Mean Scores",
    y = "Density",
    caption = "mean = 3.09, sd = 0.69, median = 3.0, skew = -0.67, kurtosis = 0.25"
  )+
  annotate("text", x = round(mean(data$trst.mean.scores), 2), y = 0.85,
           label = paste("mean =", round(mean(data$trst.mean.scores), 2)), color = "red")

# eef scores are not normally distributed, also bi-modal pos skew
eef.scores.hist <- ggplot(data, aes(eef.mean.scores))+
  geom_histogram(aes(y = after_stat(density)), bins = 15,
                 color = 1, fill = "grey")+
  geom_density(lwd = 1.5, color = 1, fill = 1, alpha = 0.35)+
  geom_segment(x = round(mean(data$eef.mean.scores), 2), y = 0, yend = 0.80, color = "red")+
  labs(
    x = "Exp. of Electoral Fraud Mean Scores",
    y = "Density",
    caption = "mean = 2.27, sd = 0.75, median = 2.2, skew = 0.19, kurtosis = -0.55"
  )+
  annotate("text", x = round(mean(data$eef.mean.scores), 2), y = 0.85,
           label = paste("mean =", round(mean(data$eef.mean.scores), 2)), color = "red")

# place plots side by side
gridExtra::grid.arrange(eef.scores.hist, trst.scores.hist, ncol = 2)
```

Given the non-normal (or at least un-equal skew of the) distribution of the two scale scores and ordinal nature of the variable items, I conducted a Spearman's rank correlation test [@spearman1907] and found a negative correlation of $\rho$ -0.49 (Kendall's \tau = -0.39) between summated scores on the two scales.

The negative correlation between the items and the composite scores makes sense. Theoretically speaking, a person who is certain that electoral fraud will occur clearly lacks trust in the election. That is, expectation of electoral fraud does imply lack of trust in elections. However, the reverse isn't also necessarily true; it is reasonable to hold a degree of trust in elections alongside expectations that electoral fraud will occur. Holding both considerations isn't necessarily an example of cognitive dissonance, as a lack of confidence in elections doesn't necessarily imply an expectation of electoral fraud. A lack of trust in elections might render one open to potentialities, but expectation of fraud is in no way necessitated by lacking trust in elections. When considered along a spectrum, both can be considered as degrees of *confidence*; and in the absence of confidence, insecure.

Accordingly, those whose responses to the "trust in elections" items would be cancelled out by their responses to the "expectation of fraud" items. Those who are, or are close to being, equally confident in both directions would be the most *insecure* about their expectations[^Insecure in terms of their confidence. Not insecure as to imply some anxious or concerned affect.]. Having no trust in elections and no expectation that electoral fraud will occur is functionally equivalent to having equal degree of both. Quite literally, an uncertain feeling of "it could go one way or the other". Operationally, this would reflect a middling 0 on a scale of *confidence*. Seeing as how expectation of electoral fraud is also a level of confidence, then it is reasonable to place all items onto a symmetric scale where positive values represent trust in elections and negative values represent expectation of electoral fraud.

The primary dependent variable is *confidence in elections*, which reflects one's trust in election administration minus their expectations of electoral fraud. Operationally, the dependent variable is a linear composite scale of item responses across items measuring trust in elections and expectations of electoral fraud re-scaled to range from -3 to 3. Items comprising trust in elections were positively coded to range from 0 "Not at all confident/committed" to 3 "Very confident/committed". Likewise, items capturing expectation of electoral fraud were coded negatively to range from 0 "Not likely at all" to -3 "Very likely".

Computed mean or sum scores for multiple Likert items are commonly acceptable, as they are relatively robust in spite of violating assumptions of a normal distribution of data (or residuals).
due to the somewhat non-normal distribution of the sum or mean scores and the ordinal nature of the variable items, an ordered regression analysis is preferred.

The primary independent variable is the prospect of military veterans as election workers (i.e., staff or volunteers).

@@

The key question is whether trust in elections (or voter confidence) is response to the treatment vignette.

Note that the question wording for the survey items derived from the 2022 SPAE are modified slightly but substantially. The question wording in this study is most similar to those in 2022 SPAE that assess confidence that votes will be counted as voters intend without specific reference to their own vote (e.g., in their county, state, or nationwide). In this study, two survey items ask about confidence in vote counts in one specific county in Arizona and another referencing their local community. Moreover, the 2022 SPAE includes "I don't know" as a response option, whereas the comparable item in this study did not include a "don't know" response option. Comparison between responses to the two similar items was conducted....

>>>>

---
title: title
subtitle: subtitle
shorttitle: A Short title # aog-article-pdf
shortauthors: Isaiah      # aog-article-pdf
authors:
  - name: Isaiah Espinoza
    email: gespinoz@umd.edu
    roles: author
    affiliation:
      - name: University of Maryland
        department: Department of Government and Politics
        city: College Park
        state: MD
        country: US
reference-section-title: References
date: "2024-10-16"
date-modified: today
knitr: 
  opts_chunk: 
    dev: "ragg_png" # same as knitr::opts_chunk$set(dev = "ragg_png")
editor_options: 
  chunk_output_type: console
---

@@

set.seed(1234)

# load packages
library(tidyverse)
library(easystats)
library(patchwork)

# load data
load("data/research-paper-df-20250203.Rdata") #df

# load data dictionary/codebook
codebook <- read.csv("data/survey_codebook.csv")
df_dict <- read.csv("data/df_dictionary.csv")

@@

That being said, the general public perception, attitudes, or even stereotypes about military veterans are significant to consider. Although military recruitment shows a downward trend as of late, public perceptions of veterans are overwhelmingly positive [@kleykamp2023]. Veterans are a particularly potent group where mention of veteran status seems to have a positive, calming, or nullifying effect on attitudes. For instance, recent research shows that, during his campaign in the 2020 Democratic Primaries, Pete Buttigieg's military background mitigated discrimination against him when he was presented as a veteran married to man [@magni2024]. Similar research has also found that a candidate's veteran status affords them better evaluations regarding competency in particular issue areas (e.g., war competence) [@hardy2019]. Moreover, veteran status seems to mitigate or nullify usual stigmas associated with mental illness. That is to say, there is negative stigma associated with mental illness [@corrigan2002] and such stigma incurs labor market discrimination [@hipes2016], but evidence suggests that veteran status overrides such stigma and discrimination [@maclean2014]. Or, in another light, mental illness is seemingly more *understandable* (i.e., permissible) for veterans given the presumptive reasons for their mental strife. And media framing as such plays an important role on public perception [@kleykamp2015].

## Theory and Hypotheses

Considering the generally positive perception afforded to veterans by the public, it is reasonable to suspect that efforts that promote recruitment of veterans and their family members to work as election staff and volunteers would boost public confidence in election administration. Indeed, that is the primary hypothesis of this study. Although citizens may vote in various ways across the country (e.g., by mail, ballot drop box, in-person), the simple announcement that election officials are engaging in efforts to recruit veterans to work as election staff may boost confidence in elections administration regardless of how, or whether, an individual plans to vote. Formally,

> H~1~: Announcements that election officials are recruiting miilitary service veterans to work as election staff and volunteers will be associated with greater confidence in elections administration compared to announcements that do not mention military veterans.

I considered that veterans would elicit high degrees of approval and support for whatever cause or issue presented. For example, I'd expect high approval and support for recruiting veterans to baseball teams just as much as for recruiting them to work as election staff. Likewise, asking participants whether they supported programs intent on recruiting anyone to work as election staff would garner support regardless of the target group. Considering that support and admiration for veterans is generally high among the population [@kleykamp2023], then discerning the impact on one's confidence would require a survey experiment designed to determine whether publicized recruitment efforts targeting veterans as a group would have any special effect on confidence in elections administration.

Since the results and events following the 2020 election loomed large in anticipation of the U.S. 2024 general elections, it is reasonable to take into account beliefs about legitimacy of the 2020 election. This group who denies the legitimacy of the 2020 election results were possibly the most distrusting of the 2024 election. However, given that veteran service members are held in high regard, generally speaking, announcing that veterans are being actively recruited to work as election staff and volunteers is expected to positively influence confidence in elections among those who remained firm in the belief that the 2020 election results were illegitimate. I take into account that legitimacy beliefs about the 2020 election and expectations of electoral fraud are related to partisanship and include partisanship as a control in the analysis.

In addition, I expect that both expectations of electoral fraud and concerns for voter safety would lessen (shrink) when presented with information that veterans are being actively recruited to work and volunteer in election offices.

Finally, prior research has shown that confidence that one's own vote would be counted as intended has been much stronger than confidence that votes nationwide would be treated likewise [@sances2015; @stewart2022]. In other words, the public is more trusting of elections closer to home. Taking into account that confidence is lower for elections conducted outside of one's local area, then information that veterans are being recruited to help conduct elections in locations beyond one's locale may close this gap between confidence in elections within one's local area and confidence in elections elsewhere. Although it may be difficult to discern whether recruiting veteran service members would improve confidence in elections near one's local area in light of such hometown favoritism, a positive effect on confidence in elections held beyond one's local area---where confidence is already expectedly lower---can help discern the impact of the group alone. Therefore, any observed disparity of confidence in elections (i.e., the "confidence gap") between elections within one's local area and elections beyond will be smaller when people are presented with information about election official recruitment efforts of veteran service members.

## Experiment Design and Survey Measures

To test the theory that publicized efforts to recruit veterans to work as election staff and volunteers would improve confidence in elections and ease insecurity, a recent experiment was embedded in a survey developed and conducted by the Center for Democracy and Civic Engagement (CDCE) at the University of Maryland. The survey was fielded from August 29th, 2024 to September 18th, 2024 on a non-probability sample of 1,287 U.S. citizens 18 years of age or older. Respondents were randomly split into either the treatment ($n =$ 650) control ($n =$ 637) conditions[^4].

Survey participants read either a treatment or control vignette, which was a fabricated news article about efforts in Maricopa County, AZ to recruit election staff and volunteers for the 2024 general election. The treatment vignette referred to a program designed to recruit veterans and their family members and describes an interviewee "Jordan Braxton" as an Army veteran. The control vignette simply omitted any mention of veterans and their family members, and didn't describe "Jordan Braxton" as an Army veteran. Beyond those small differences and the headlines, the article vignettes are identical. Therefore, effects can be attributed to the information about veterans in the treatment vignette[^5].

It should be noted that Maricopa County, AZ was chosen as the setting of the story in the vignette due to the increased scrutiny levied toward election administration there after the 2020 election [@giles2021; @maricopacountyelectionsdepartment2022]. Because of this, treatment effects are potentially limited or constrained to attitudes concerning the location specified in the vignette. A block of survey items asked specifically about Maricopa County, AZ, followed by an identical block of items that asked the same questions about one's local area. Later, I compare and discuss results between the items pertaining to different settings.

@@

Note that, except for the item assessing voter confidence (i.e., votes counted as voters intend), most other question items from the 2022 SPAE inquire into the voter experience after the election has occurred. The items in this study, however, all inquire into confidence over one's expectations in anticipation of upcoming elections. In addition, other survey questionnaires ask similar questions that specifically contrast between a respondent's own local community, their state, and nationwide; this is in addition, or complementary to, items that distinguish between local area officials, state election officials, state government, and the U.S. federal government. The survey of this study, instead, focused on attitudes regarding election workers such as election officials, staff, and volunteers specific to the location mentioned in the vignette---Maricopa County, AZ---as well as specific to one's local area[^6].

The survey included items that assessed both trust in elections administration and expectations of electoral fraud, referred to simply as *trust* and *distrust*, respectively. Responses from these two sets of items were utilized to construct the primary dependent variable, *confidence in elections*.

Prior to the reading the treatment or control vignette, participants responded to distinct survey items assessing one's favorability to local election officials ("In general, how favorable or unfavorable is your impression of local election officials?") and legitimacy of the 2020 election results ("Regardless of whom you supported in the 2020 election, do you think Joe Biden's election as president was legitimate, or was he not legitimately elected?"). The former presented five response options from "Strongly favorable" to "Strongly unfavorable"; the latter item presented a binary of either "Legitimate" or "Not Legitimate".

The conventional multiple item assessment of partisanship was included and used to construct a variable of partisanship consisting of three categories: Democrat, Republican, and Independent. In this case, Independents are "true" Independents in that they do not identify with either political party and do not "lean" toward either Republican or Democratic parties. Self-described "Independent leaners" are re-coded as falling into the partisan category in which they lean e.g., Independents who expressly report that they lean Republican are denoted as Republican.

Finally, a pair of survey items assessed concerns for potential violence and confidence in voter safety. The pair was asked once in regard to Maricopa County, AZ and again with respect to one's local area[^7].

### Trust and Distrust in Elections

A set of five survey items measured public trust in elections by inquiring into the degree of confidence that votes will be counted as voters intend, confidence that electoral systems are secure from technological threats, perceived commitment of election staff, confidence that outcomes will be fair, and confidence that the voting process will be fair. Each item in this series presented four response options from "Not at all \[confident/committed\]", "Not too \[confident/committed\]", "Somewhat \[confident/committed\]", and "Very \[confident/committed\]". In total, five survey items measure different aspects of trust in elections administration. The items measuring confidence in elections include:

Another series of five survey items captured an individual's level of *distrust* in elections administration based on the extent to which they *expect electoral fraud* to occur. These items were prefaced with the question, "How likely do you think any or all of the following will happen during this year´s elections in \[Maricopa County, AZ/ your local area\]?" Each item presented four response options: "Not likely at all", "Not too likely", "Somewhat likely", and "Very likely".

```{r}
#| label: tbl-items
#| tbl-cap: "Survey Items for Trust and Distrust in Elections"

trust_txt <- df |> select(q19, q20, q21, q22, q24) |>
  labelled::set_variable_labels(
    q21 = "How commited do you think election staff and volunteers in Maricopa County, AZ will be to making sure the elections are fair and accurate?"
  ) |> 
  labelled::generate_dictionary() |> 
  dplyr::as_tibble() |> 
  select(Trust=label)

distrust_txt <- df |> select(q28_1:q28_5) |>
  labelled::generate_dictionary() |> 
  dplyr::as_tibble() |> 
  select(Distrust=label)

conf_items_text <- trust_txt |> 
  mutate(Distrust = pull(distrust_txt))

kableExtra::kbl(conf_items_text, booktabs = T,
                col.names = c("Trust", "Distrust"),
                linesep = "\\addlinespace") |> 
  kableExtra::kable_styling(full_width = T,
                            font_size = 8,
                            latex_options = c("HOLD_position")) |> 
  kableExtra::column_spec(column = 1:2, width = "8cm")

Similar to the items capturing trust in elections (@tbl-items), I computed a summated score for each respondent reflecting the extent of their distrust in elections.

Upon construction of the two summated scales of *trust* and *distrust*, internal consistency reliability coefficient $\alpha$ were estimated for both and found to be adequate ($\alpha$ \> 0.8); the trust in elections scale $\alpha$ = 0.93 and the distrust scale $\alpha$ = 0.86. In other words, the items in each of the scales were intercorrelated well enough for each to compose a single scale presuming a unidimensional construct for each.

I expected item responses from these two scales to be inversely correlated among the sample, but not mutually exclusive. Indeed, this is what I find. Polychoric correlations between the trust and distrust items negatively correlate as expected\[\^See Appendix C\], as does the correlation between the two sum score scales. Given the ordinal nature of the variable items, I conducted a Spearman's rank correlation test [@spearman1907] and found a negative correlation of $\rho = -0.49$ ($95\%$ CI \[-0.53, -0.44\]; Kendall's $\tau = -0.39$) between scores on the two scales for the AZ items (For the local item scale score correlation: $\rho = -0.52$, $95\%$ CI \[-0.56, -0.48\]; Kendall's $\tau = -0.39$). The negative correlation between the trust and distrust items and scores makes intuitive sense.

@@

The primary outcome of interest concerns public confidence in elections administration compensated by one's distrust in elections, i.e., their expectations of electoral fraud. Because trust and distrust are assessed by multiple Likert items, conventional methods call for combining the items of each set into separate Likert scales (also known as summated rating scales). Since the item response options are not symmetric---i.e., not bipolar with contrasting positive and negative response options to include a neutral middle option---then the *trust in elections* scale could only be understood as a measure of trust in the positive use of the term (i.e., unipolar). Likewise, the expectation of fraud items fit to a scale that corresponds to a unipolar measure of confident *distrust*, so to speak. However, many of the items comprising the trust scale directly oppose items capturing distrust. For instance, one item from the trust scale has the respondent report their confidence that votes will be counted as voter's intend, whereas a separate item from the distrust scale has the the respondent report their expectation that many votes will not be counted. Positive responses to these two items directly contradict each other, and would thus cancel out.

The relationship is that trust and distrust represent opposing degrees of confidence along the same spectrum. However, despite the inherent contrast, one can hold both positive degrees of trust and distrust, feeling similarly confident in both respects. Accordingly, equivalent degrees of trust and distrust cancel out, rendering one relatively more *insecure* about their expectations of the future despite the fact that they may feel confident in either direction. When trust and distrust are considered as reflective of one's confidence about future expectations, then both can be placed along the same spectrum. Therefore, a lack of confidence denotes insecurity.

Something may cause a person to have greater expectations that election fraud will occur, which in turn will always lower their trust in elections administration. In contrast, a person may feel a bit more confident in the integrity of elections, but this confidence is constrained by the extent to which they expect election fraud to occur. This means that a person's baseline level of confidence in elections admin must take into account their present degree of distrust---i.e., their expectations of electoral fraud.

Those who are, or are close to being, equally confident in both directions would be the most *insecure* about their expectations. Having no trust in elections and no expectation that electoral fraud will occur is functionally equivalent to having equal degree of both. Quite literally, this reflects an uncertain feeling of "it could go one way or the other" with respect to the integrity of elections administration. Operationally, this would reflect a middling point on a scale of *confidence*.

In light of this, I reverse coded positive expectations of electoral fraud and added responses from the distrust scale to the trust in elections scale. That is, item responses on the trust scale ranged from 0 ("Not confident at all") to 3 ("Very confident), whereas responses on the distrust scale ranged from 0 ("Not likely at all") to -3 ("Very likely"). Combining responses in this way resulted in an overall measure of *confidence in elections* to serve as the dependent variable[^8]. On this scale, positive values reflect higher confidence in elections, negative values reflect greater distrust, and values closer to zero correspond to an individual's degree of insecurity.

The advantage is that this permits a better test of the treatment effect as theorized. Of course, it is easy enough to examine the effect of the treatment on both trust and distrust separately. However, the theory presents a hypothesis that the treatment will impose influence upon both variables *concurrently and in a positive direction* with respect to the relationship between trust and distrust. That is to say, confidence increases insofar as trust in elections increases and distrust decreases; since distrust detracts from trust, then an accurate measure of confidence must take that constraint into account.

@@

I first review responses to the survey items capturing trust and distrust separately by treatment condition. When distinguished by experiment condition, trust in elections admin is noticeably higher for those who read the treatment vignette. The difference is more pronounced for some items over others. That being said, there is a definite distinction between items pertaining to Maricopa County, AZ compared to items that pertain to a respondent's local area. In fact, the treatment appears to have hardly made a difference when the items concerned one's local area.

```{r}
#| label: fig-trustlikert
#| fig-cap: "Trust in Election Administration by Experiment Condition"
#| fig-height: 4.5
#| fig-pos: "h"

# Likert plot of trust and confidence in election admin comparing treatment and
# control groups
trust.likert.treat.az <- df |>
  # relabel factor labels for plotting
  mutate(across(
    c(q19, q20, q21, q22, q23, q24, q30, q31, q32, q33, q34, q35),
    ~ pewmethods::fct_case_when(
      .x %in% c("Not at all confident", "Not at all committed") ~ "Not at all confident",
      .x %in% c("Not too confident", "Not too committed") ~ "Not too confident",
      .x %in% c("Somewhat confident", "Somewhat committed") ~ "Somewhat confident",
      .x %in% c("Very confident", "Very committed") ~ "Very confident",
      TRUE ~ NA
    )
  )) |>
  ggstats::gglikert(
    include = c(q19, q20, q21, q22, q24),
    y = 'group',
    facet_rows = vars(.question),
    symmetric = F,
    labels_size = 2.5,
    add_totals = T
  ) +
  ggplot2::scale_fill_grey(start = 0.5, end = 0.1)+
  # ggplot2::scale_fill_brewer(palette = "Greys")+
  labs(subtitle = str_wrap("Maricopa County, Arizona", width = 95)) +
  theme_bw() +
  theme(
    text = element_text(family = "Palatino Linotype"),
    legend.position = 'bottom',
    axis.text.x = element_blank(), # remove percentage text along x-axis
    strip.text.y.right = element_blank()
  )

# confidence in elections between experiment conditions for items concerning local area
trust.likert.treat.lcl <- df |>
  # relabel factor labels for plotting
  mutate(across(
    c(q19, q20, q21, q22, q23, q24, q30, q31, q32, q33, q34, q35),
    ~ pewmethods::fct_case_when(
      .x %in% c("Not at all confident", "Not at all committed") ~ "Not at all confident",
      .x %in% c("Not too confident", "Not too committed") ~ "Not too confident",
      .x %in% c("Somewhat confident", "Somewhat committed") ~ "Somewhat confident",
      .x %in% c("Very confident", "Very committed") ~ "Very confident",
      TRUE ~ NA
    )
  )) |>
  ggstats::gglikert(
    include = c(q30, q31, q32, q33, q35),
    y = 'group',
    facet_rows = vars(.question),
    variable_labels = c(
      q30 = "Votes counted as intended",
      q31 = "Election officials will do good job",
      q32 = "Election workers will be committed",
      q33 = "Voting process will be fair",
      q35 = "Election technology will be secure"
    ),
    facet_label_wrap = 10,
    labels_size = 2.5,
    symmetric = F,
    add_totals = T
  ) +
  ggplot2::scale_fill_grey(start = 0.5, end = 0.1)+
  # ggplot2::scale_fill_brewer(palette = "Greys")+
  labs(subtitle = str_wrap("Local Area", width = 95)) +
  theme_bw() +
  theme(
    text = element_text(family = "Palatino Linotype"),
    legend.position = 'bottom',
    axis.text.y.left = element_blank(), # remove y-axis text on 2nd plot
    axis.text.x = element_blank(), # remove percentage text along x-axis
    strip.text.y.right = element_text(angle = 0, size = 8)
  )

# trust.likert.treat.az
# trust.likert.treat.lcl

# Instead of two separate plots, arrange into one plot while remaining separated
# use `ggpubr` package to print multiple plots side-by-side (or alt. row-by-row)
ggpubr::ggarrange(
  trust.likert.treat.az,
  trust.likert.treat.lcl,
  ncol = 2,
  common.legend = T,
  legend = "bottom", 
  font.label = list(family = "Palatino Linotype")
) +
  # add figure footnote annotation
  patchwork::plot_annotation(
    caption = str_wrap("Missing values removed. Displays percent proportions of responses by experiment condition and combined percentage proportions for two categories.",
               width = 85), theme = theme(text = element_text(family = "Palatino Linotype")))

Reviewing item responses distinguished by treatment condition reveals particulars of the treatment effect (@fig-trustlikert). Among the items that asked about Maricopa County, AZ, it is a bit easier to see upon which items the treatment vignette had the greatest effect. It appears that the main effect of the treatment is upon responses to the item that measures confidence in the security of election technology.

Regarding confidence in the commitment of election staff and volunteers, a higher proportion of respondents in the treatment group selected "Very confident" over "Somewhat confident" compared to the control group, but only for items concerning Maricopa County, AZ. Although confidence in election workers was generally high across the board, the shift from "Somewhat" to "Very" confident for items pertaining to election worker *commitment* to fairness and accuracy seems to clearly speak to the group being recruited, i.e., veterans and family members. However, the same visible shift in proportion from "Somewhat" to "Very" confident isn't observed when items asked about one's local area, which undermines the suggestion that the treatment effect can be cleanly attributed to respondent's feelings about veteran service members and their families. It may be the case that respondents harbor particular attitudes or sentiment about Maricopa County, AZ itself, or there may be nothing uniquely special about Maricopa County, AZ supposing that elections anywhere beyond one's local area generally garner slightly less confidence.

What this does suggest, however, is that trust in the fairness and accuracy of the electoral process is notably higher upon the notion that military veterans will be directly engaged as election staff and volunteers in places where such trust falters. The question turns to whether this also lowers distrust, or the extent to which one expects electoral fraud to occur.

```{r}
#| label: fig-distrustlikert
#| fig-cap: "Disrust in Election Administration by Experiment Condition"
#| fig-height: 4.5
#| fig-pos: "h"

# distrust in elections in AZ by experimental condition
distrust.likert.az <- df |>
  # reverse levels of factor vars to show positive difference in "not likely"
  mutate(across(c(q28_1:q28_5), ~forcats::fct_rev(.))) |> 
  ggstats::gglikert(
    include = c(q28_1:q28_5),
    y = 'group',
    facet_rows = vars(.question),
    variable_labels = c(
      q28_1 = "Voter fraud will occur",
      q28_2 = "Many votes won't be counted",
      q28_3 = "Many voters will be turned away",
      q28_4 = "Foreign Interference",
      q28_5 = "Election officials will discourage voting"),
    facet_label_wrap = 10,
    labels_size = 2.5,
    symmetric = F,
    add_totals = T
  ) +
  ggplot2::scale_fill_grey(start = 0.5, end = 0.1)+
  # ggplot2::scale_fill_brewer(palette = "Greys")+
  labs(subtitle = str_wrap("Maricopa County, Arizona", width = 95)) +
  theme_bw() +
  theme(
    text = element_text(family = "Palatino Linotype"),
    legend.position = 'bottom',
    axis.text.x = element_blank(), # remove percentage text along x-axis
    strip.text.y.right = element_blank()
  )
